\subsubsection{Libraries providing regular expression functionality}
For most popular programming languages, various regular expression functions are provided using standard libraries or are built into the language.  Table~\ref{table:languagesLibrariesStandards} describes the standard regular expression libraries or built-ins \emph{provided as a core language feature} for all but three of the top 20 most popular languages ordered according to the TIOBE\footurl{http://www.tiobe.com/tiobe_index} index. The C, Visual Basic and Assembly Language languages (ranks 2, 12 and 13, respectively) do not provide built-in regular expression support and so are not shown.

\input{table/languagesLibrariesStandards}

Although pure ANSI C does not include a standard regex library or built-in, libraries providing regex support can be made available such as \underline{POSIX}, \underline{PCRE} or \underline{re2c}.  Similarly, pure Visual Basic has no core regex support but can use the \underline{RegExp} object provided by the VBScript library.  In fact, for most general-purpose languages, multiple alternative regular expression libraries can be found which may offer slightly different syntax or optimizations for speed.  Some of these libraries implement a language defined by a standard (like PCRE or ECMAScript) or offer a choice of languages.  For example, the std::regex library\footurl{http://en.cppreference.com/w/cpp/regex/syntax_option_type} implements engines for ECMAScript Regular Expressions (the default), AWK Regular Expressions,  POSIX BRE or POSIX ERE.  The following libraries are alternatives to std::regex: Boost.Regex, Boost.Xpressive, cppre, DEELX, GRETA, Qt/QRegExp and RE2.  These alternative libraries are developed by hobby users and software giants alike, with RE2~\cite{re2} being a recent and notable alternative library developed by Google.
