\section{History of regex}

\subsection{Kleene's theory of regular events}
In 1943 a model for how nets of nerves might `reason' to react to patterns of stimulus was proposed in a paper by McCulloch-Pitts.  In 1951, Kleene further developed this model with the idea of 'regular events'.  In his terminology, `events' are all inputs on a set of neurons in discrete time, a `definite event' E is some explicit sequence of events, and a `regular event' is defined using three operators: 1. logical or, 2. concatenation and 3. the Kleene star which represents zero or more of some definite event.  Kleene showed that `all and only regular events can be represented by nerve nets or finite automata', and went on to show that operations on regular events are closed, and to define an algebra for simplifying regular events.  The formulas used to describe regular events were named `regular expressions' in Kleene's 1956 refined paper.

\subsection{First regex compiler}
Many additional formalisms were built on Kleene's set of three operators, and then around 1967\footurl{https://www.bell-labs.com/usr/dmr/www/qed.html}, Ken Thompson implemented the first regular expression compiler in IBM 7090 assembly for a version of `qed' (quick editor) at Bell labs.  Existing editors were only able to search/replace using whole words.  Thompson's editor was able to search and replace using the features STR, END, ANY, CCC, NCCC and KLE\footurl{https://www.bell-labs.com/usr/dmr/www/qedman.html}, which provide a language equivalent to Kleene's regular expressions.

\subsection{Early regex in Unix}
Thompson went on to create Unix in 1969 with Dennis M. Ritchie, the core of which was an assembler, a shell and `ed' - an editor with regex search/replace capabilities based on qed.  Unix tools grep (1973), sed (1974) and awk (1977) notably relied on regular expressions.  The feature set of regular expressions evolved over time, and although it is outside the scope of this thesis to capture all details of this evolutionary process, the next major milestone is the release of egrep and the extended regular expressions (POSIX ERE) standard in 1979?, which adds the features CG, BKR, SNG, DBB, LWB, QST, ADD and OR as well as 12 default character classes similar to DEC, WRD, WSP, NDEC, NWRD and NWSP but using syntax like \verb![:digit:]! instead of the modern \verb!\d! for DEC.  The BKR feature is noteworthy in that it is the first feature to extend the set of languages expressible by regex beyond the regular languages described by Kleene as described in Section\todoMid{N}.

\subsection{Maturity of standards}
In 1979, Hopcroft and Ullman published the `Cindarella' textbook covering automata and theory behind regular expressions in Unix (excluding back-references).  Perl 2 was released in 1988 with regex support including modern default character classes like \verb!\d! for DEC, significantly boosting the popularity and user base of regex.  In 1992, Henry Spencer released {\tt regcomp}, a major regex library for C.  In the same year, the POSIX.2 standard was also released.  This standard includes specifications for regex features, although it was not ultimately adhered to by many language developers.  In 1997, the O'Reilly book `Mastering Regular Expressions' was first published, and the PCRE standard was first released.  In 2000, Python 2 was released including the `Secret Labs' {\tt re} module\footurl{https://github.com/python/cpython/commits/master/Lib/re.py?page=2} which implemented named capture groups and began including Unicode support in regex.
